---
title: "Data Catalogue"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data Catalogue}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  eval = FALSE,
  comment = "#>"
)
```

## Introduction

`fugazi.db` ships six lazy-loaded data objects - the primary source data
behind the Fugazi Live Series. This vignette describes each table and, most
importantly, **how they join together** - see each object's own help page
(e.g. `?fls_data`) for the full column-by-column description.

This vignette contains no runnable code of its own - `fugazi.db` has no
functions, so every example below uses
[`Repeatr`](https://github.com/alexmitrani/Repeatr), which depends on
`fugazi.db` for its raw inputs. Install both to run these examples.

## The tables

| Object | One row per | Primary/join keys |
|---|---|---|
| `fls_data` | show | `gid`; `country`+`city`+`venue` |
| `fls_tags_raw` | tagged MP3 track | `album` (encodes date/venue/city/subdivision/country as text - not a direct `gid` join, see below) |
| `songvarslookup` | song | `song` (title text); `releaseid` |
| `releases` | release (album/EP) | `releaseid` |
| `song_tempo_bpm_data` | song | `song` (title text) |
| `fls_venue_geocoding` | venue | `country`+`city`+`venue` |

## Join keys

### `gid` - the show identifier

`gid` is a slug built from city, country, and date (e.g.
`washington-dc-usa-90387`), and is the primary key of `fls_data`. It is also
the primary key of every show-level object `Repeatr` derives from `fls_data`
(`othervariables`, `shows_data`, and others) - join on `gid` to attach any of
those to `fls_data` or to each other.

```{r}
library(fugazi.db)
library(dplyr)

# every show, with its Repeatr-derived sound quality rating
fls_data %>%
  left_join(Repeatr::gid_sound_quality, by = "gid")
```

`fls_tags_raw` does **not** carry `gid` directly - its `album` field encodes
a show's date and venue as free text (`"YYYYMMDD Venue, City, State,
Country"`). `Repeatr::Repeatr_1()` is the authoritative parser for this
field (matching it back to `gid` via the show data, not by splitting on
commas, since a venue or city name can itself contain a comma); don't
re-parse `album` by hand.

### `country` + `city` + `venue` - venue coordinates

`fls_venue_geocoding` is keyed by the combination of `country`, `city`, and
`venue` (with `city` disambiguated as `"City (ST)"`/`"City (Country)"` for
towns that share a name with another tour stop - Portland, Columbia,
Croydon, Oxford, Newcastle). `fls_data`'s own `country`/`city`/`venue`
columns use the *undisambiguated* city form; `Repeatr::Repeatr_1()` applies
the same disambiguation rule before joining against `fls_venue_geocoding`.

```{r}
# venues with no confirmed coordinates yet
fls_venue_geocoding %>% filter(is.na(x) | is.na(y))
```

### `song` - discography metadata

`songvarslookup` and `song_tempo_bpm_data` are both keyed by `song` (the
song title, as text - not a numeric id). `Repeatr` assigns a stable numeric
`songid` once it classifies the full song set (see `Repeatr::songidlookup`);
that id doesn't exist in `fugazi.db`, precisely so these two files can't
silently drift out of sync with `Repeatr`'s classification logic by
hardcoding an id that might not match.

```{r}
# every song's tempo alongside its discography metadata
songvarslookup %>%
  left_join(song_tempo_bpm_data, by = "song")
```

### `releaseid` - discography structure

`releases` is keyed by `releaseid`, referenced by `songvarslookup$releaseid`.

```{r}
# every song, with its release's name and colour code
songvarslookup %>%
  left_join(releases, by = "releaseid") %>%
  select(song, release, colour_code)
```

## Where the derived data lives

`fugazi.db` intentionally stops at primary data. Song classification,
choice-model ratings, tour/attendance summaries, and everything else
`Repeatr` computes from these tables lives in `Repeatr` itself - see its
`vignette("Data-Provenance")` for the full catalogue of derived objects, and
`vignette("Rebuilding-the-Data")` for how `fugazi.db`'s own `data-raw/`
sources get refreshed and turned into the objects documented here.
