---
title: "Data Catalogue"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data Catalogue}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  eval = FALSE,
  comment = "#>"
)
```

## Introduction

`fugazibase` includes six lazy-loaded data objects - cleaned, corrected
reference data behind the Fugazi Live Series. This vignette describes each
table and **how they can be joined together**. See each object's
own help page (e.g. `?shows`) for the full column-by-column description.

This vignette contains no runnable code of its own - `fugazibase` has no
functions, so every example below uses only base R and dplyr on this package's
own data. Each table has been cleaned, corrected, and reformatted from the sources described in the package README and each table's own help page (e.g. `?shows`), and is updated occasionally as new information becomes available.

## The tables

| Object | One row per | Primary/join keys |
|---|---|---|
| `shows` | show | `gid`; `country`+`city`+`venue` |
| `locations` | venue | `country`+`city`+`venue` |
| `durations` | tagged MP3 track | `gid`; `song` |
| `discography` | release (album/EP) | `releaseid` |
| `songs` | song | `song` (title); `releaseid` |
| `bands` | show + band | `gid` |

## Joining tables

### `gid` - the show identifier

`gid` (gig id) is a slug built from city, country, and date (e.g.
`washington-dc-usa-90387`), and is the primary key of `shows`. It is
also the primary key of `durations` and `bands` - either of those
can be joined to `shows` using `gid`.

```{r}
library(fugazibase)
library(dplyr)

# every show with its revorded tracks
shows_tracks <- shows |>
  left_join(durations, by = "gid")

shows_tracks
```

### `country` + `city` + `venue` - venue coordinates

`locations` is keyed by the combination of `country`, `city`, and
`venue` - join it onto `shows` on those three columns to attach
coordinates to a show. `city` is the plain city name, the same as in
`shows`; a handful of towns share a name with another tour stop
(Portland, Columbia, Croydon, Newcastle, Oxford, Springfield), so
`country` and `venue` are both needed alongside `city` to pin down a
single row. Coordinates are latitude/longitude in decimal degrees, using
the WGS 84 datum. 

```{r}
# every show, with its venue's coordinates
shows_locations <- shows |>
  left_join(locations, by = c("country", "city", "venue"))

shows_locations
```

### `song` - song metadata

`songs` is keyed by `song` (the song title, as text), and carries
studio discography metadata (release, vocals, instrumental, duration).
`durations` is also keyed by `song` text for each tagged track - join the two
to attach song metadata to a specific live rendition.

```{r}
# every recorded track, with song information
durations_songs <- durations |>
  left_join(songs, by = "song")

durations_songs
```

### `releaseid` - discography structure

`discography` is keyed by `releaseid`, referenced by `songs$releaseid`.

```{r}
# every song, with its release's name and date
songs_discography <- songs |>
  left_join(discography, by = "releaseid") |>
  select(song, release_track, release, releasedate) |>
  arrange(releasedate, release_track)

songs_discography
```

This package's data is refreshed periodically by the maintainer as new
shows are documented or corrections are made to the underlying sources.
