---
title: "Data Catalogue"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data Catalogue}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  eval = FALSE,
  comment = "#>"
)
```

## Introduction

`fugazi.db` ships nine lazy-loaded data objects - cleaned, corrected
reference data behind the Fugazi Live Series. This vignette describes each
table and, most importantly, **how they join together** - see each object's
own help page (e.g. `?fls_shows`) for the full column-by-column description.

This vignette contains no runnable code of its own - `fugazi.db` has no
functions, so every example below uses only base R/dplyr on this package's
own data. Every table here is generated by
[`Repeatr`](https://github.com/alexmitrani/Repeatr)'s
`export_fugazidb_data()`, which holds all the cleaning, correction, and
classification logic - see its `vignette("Data-Provenance")` for how each
table is derived.

## The tables

| Object | One row per | Primary/join keys |
|---|---|---|
| `fls_shows` | show | `gid`; `country`+`city`+`venue` |
| `fls_venue_geocoding` | venue | `country`+`city`+`venue` |
| `fls_tags` | tagged MP3 track | `gid` |
| `releases` | release (album/EP) | `releaseid` |
| `songvarslookup` | song | `song` (title text); `releaseid` |
| `song_tempo_bpm_data` | song | `song` (title text) |
| `songidlookup` | song | `songid`; `song` (title text) |
| `played_with` | show + co-billed band | `gid` |
| `played_with_data` | show + co-billed band | `gid` |

## Join keys

### `gid` - the show identifier

`gid` is a slug built from city, country, and date (e.g.
`washington-dc-usa-90387`), and is the primary key of `fls_shows`. It is
also the primary key of `fls_tags`, `played_with`, and `played_with_data` -
join any of those to `fls_shows` (or to each other) on `gid`.

```{r}
library(fugazi.db)
library(dplyr)

# every show, with its tagged tracks
fls_shows %>%
  left_join(fls_tags, by = "gid")
```

### `country` + `city` + `venue` - venue coordinates

`fls_venue_geocoding` is keyed by the combination of `country`, `city`, and
`venue` (with `city` disambiguated as `"City (ST)"`/`"City (Country)"` for
towns that share a name with another tour stop - Portland, Columbia,
Croydon, Oxford, Newcastle). `fls_shows` already carries its own resolved
`x`/`y` coordinates directly (joined in by `Repeatr::Repeatr_1()`, including
a handful of hardcoded per-venue corrections not present in
`fls_venue_geocoding` itself) - `fls_venue_geocoding` is useful on its own
for auditing or extending venue coverage, not required to get coordinates
onto `fls_shows`.

```{r}
# venues with no confirmed coordinates yet
fls_venue_geocoding %>% filter(is.na(x) | is.na(y))
```

### `song` - discography metadata

`songvarslookup` and `song_tempo_bpm_data` are both keyed by `song` (the
song title, as text). `songidlookup` additionally provides a stable numeric
`songid` for every song `Repeatr` recognizes - join on `song` text to attach
it to either of the other two.

```{r}
# every song's tempo alongside its discography metadata and stable songid
songvarslookup %>%
  left_join(song_tempo_bpm_data, by = "song") %>%
  left_join(songidlookup, by = "song")
```

### `releaseid` - discography structure

`releases` is keyed by `releaseid`, referenced by `songvarslookup$releaseid`.

```{r}
# every song, with its release's name and date
songvarslookup %>%
  left_join(releases, by = "releaseid") %>%
  select(song, release, releasedate)
```

## What's excluded, and why

`fugazi.db` ships facts - corrected and reformatted, but not joined,
summarized, or modelled. It deliberately excludes:

- **Free-text show notes** (`fls_notes` in Repeatr's own raw data) - a
  copyright precaution, since that text is Dischord Records' own writing
  about each show, not a structured fact like a date or venue name. See this
  package's `LICENSE`.
- **Joined/summarized/modelled data** - tour-level and song-level summaries,
  cumulative counts, transition matrices, sound-quality "stacks," and the
  `mlogit` choice model and everything derived from it. All of that lives in
  `Repeatr` itself - see its `vignette("Data-Provenance")` for the full
  catalogue of derived objects.
- **Song classification detail** - `Repeatr`'s song-title recoding and
  `tracktype` rules involve judgment calls that can shift which songs exist
  and how they're numbered. `songidlookup` ships the resulting `songid` as
  pure join-key infrastructure, but not the classification rules themselves.

See `Repeatr::export_fugazidb_data()` and its `vignette("Rebuilding-the-Data")`
for how this package's own `data-raw/` sources get refreshed.
