---
title: "Data Catalogue"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data Catalogue}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  eval = FALSE,
  comment = "#>"
)
```

## Introduction

`fugazi.db` ships six lazy-loaded data objects - cleaned, corrected
reference data behind the Fugazi Live Series. This vignette describes each
table and, most importantly, **how they join together** - see each object's
own help page (e.g. `?fls_shows`) for the full column-by-column description.

This vignette contains no runnable code of its own - `fugazi.db` has no
functions, so every example below uses only base R/dplyr on this package's
own data. Every table here is generated by
[`Repeatr`](https://github.com/alexmitrani/Repeatr)'s
`export_fugazidb_data()`, which holds all the cleaning, correction, and
classification logic - see its `vignette("Data-Provenance")` for how each
table is derived.

## The tables

| Object | One row per | Primary/join keys |
|---|---|---|
| `fls_shows` | show | `gid`; `country`+`city`+`venue` |
| `fls_venue_geocoding` | venue | `country`+`city`+`venue` |
| `fls_tags` | tagged MP3 track | `gid`; `song` |
| `releases` | release (album/EP) | `releaseid` |
| `discography` | song | `song` (title text); `releaseid` |
| `played_with` | show + co-billed band | `gid` |

## Join keys

### `gid` - the show identifier

`gid` is a slug built from city, country, and date (e.g.
`washington-dc-usa-90387`), and is the primary key of `fls_shows`. It is
also the primary key of `fls_tags` and `played_with` - join either of those
to `fls_shows` (or to each other) on `gid`.

```{r}
library(fugazi.db)
library(dplyr)

# every show, with its tagged tracks
fls_shows %>%
  left_join(fls_tags, by = "gid")
```

### `country` + `city` + `venue` - venue coordinates

`fls_venue_geocoding` is keyed by the combination of `country`, `city`, and
`venue` (with `city` disambiguated as `"City (ST)"`/`"City (Country)"` for
towns that share a name with another tour stop - Portland, Columbia,
Croydon, Oxford, Newcastle) - join it onto `fls_shows` on those three
columns to attach coordinates to a show. Note that `Repeatr::Repeatr_1()`
resolves a handful of venues via hardcoded per-venue corrections not present
in this table, so this join won't cover every show.

```{r}
# every show, with its venue's coordinates
fls_shows %>%
  left_join(fls_venue_geocoding, by = c("country", "city", "venue"))

# venues with no confirmed coordinates yet
fls_venue_geocoding %>% filter(is.na(x) | is.na(y))
```

### `song` - discography metadata

`discography` is keyed by `song` (the song title, as text), and carries
studio discography metadata (release, vocals, instrumental, duration).
`fls_tags` is also keyed by `song` text for each tagged track - join the two
to attach discography metadata to a specific live rendition.

```{r}
# every tagged track, with its discography metadata
fls_tags %>%
  left_join(discography, by = "song")
```

### `releaseid` - discography structure

`releases` is keyed by `releaseid`, referenced by `discography$releaseid`.

```{r}
# every song, with its release's name and date
discography %>%
  left_join(releases, by = "releaseid") %>%
  select(song, release, releasedate)
```

See `Repeatr::export_fugazidb_data()` and its `vignette("Rebuilding-the-Data")`
for how this package's data gets refreshed.
