---
title: "Data Catalogue"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data Catalogue}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  eval = FALSE,
  comment = "#>"
)
```

## Introduction

`fugazibase` includes six tables of data from the Fugazi Live Series. 
This vignette describes each table and **how they can be joined together**. 
See each object's own help page (e.g. `?shows`) for the full column-by-column description.

This vignette contains no runnable code of its own - `fugazibase` has no
functions, so every example below uses only base R and dplyr on this package's
own data. Each table has been cleaned, corrected, and reformatted from the sources described in the package README the corresponding help page (e.g. `?shows`).

## The tables

| Object | One row per | Primary/join keys |
|---|---|---|
| `shows` | show | `gid`; `country`+`city`+`venue` |
| `locations` | venue | `country`+`city`+`venue` |
| `durations` | track | `gid`; `title` |
| `discography` | release | `rid` |
| `songs` | song | `title`; `rid` |
| `bands` | show + band | `gid` |

## Joining tables

### `gid` - the show identifier

`gid` (gig id) is a slug built from city, country, and date (e.g.
`washington-dc-usa-90387`), and is the primary key of `shows`. 
It is also included in `durations` and `bands`, both of which
can be joined to `shows` using `gid`.

```{r}
library(fugazibase)
library(dplyr)

# every show with its revorded tracks
shows_tracks <- shows |>
  left_join(durations, by = "gid")

shows_tracks
```

### `country` + `city` + `venue` - venue coordinates

The observations in `locations` are identified by the combination 
of `country`, `city`, and `venue`. Locations data can be added to `shows` 
by joining on those three columns. `city` is the city name, the same as in
`shows`; a handful of towns share a name with another tour stop
(Portland, Columbia, Croydon, Newcastle, Oxford, Springfield), so
`country` and `venue` are both needed together with `city` to pin down a
specific location. Coordinates are latitude and longitude in decimal degrees, 
using the WGS 84 datum. 

```{r}
# every show, with its venue's coordinates
shows_locations <- shows |>
  left_join(locations, by = c("country", "city", "venue"))

shows_locations
```

### `title` - songs

The observations in `songs` are identified by `title` (the song title, as text), and include studio discography information (release, vocals, instrumental, duration).
`durations` also includes `title` for each track - join the two tables
to attach song metadata to the corresponding live renditions.

```{r}
# every recorded track, with song information
durations_songs <- durations |>
  left_join(songs, by = "title")

durations_songs
```

### `rid` - discography structure

`discography` is keyed by `rid` (release id), which is also included in `songs`.

```{r}
# every song, with its release's name and date
songs_discography <- songs |>
  left_join(discography, by = "rid") |>
  select(title, release_track, release_title, release_date) |>
  arrange(release_date, release_track)

songs_discography
```

