---
title: "Data Catalogue"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data Catalogue}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  eval = FALSE,
  comment = "#>"
)
```

## Introduction

`fugazi.db` ships six lazy-loaded data objects - cleaned, corrected
reference data behind the Fugazi Live Series. This vignette describes each
table and, most importantly, **how they join together** - see each object's
own help page (e.g. `?shows`) for the full column-by-column description.

This vignette contains no runnable code of its own - `fugazi.db` has no
functions, so every example below uses only base R/dplyr on this package's
own data. Each table is cleaned, corrected, and reformatted by the package
maintainer from the sources described in the package README and each
table's own help page (e.g. `?shows`), and is refreshed periodically as new
shows are documented or corrections are made to the underlying sources.

## The tables

| Object | One row per | Primary/join keys |
|---|---|---|
| `shows` | show | `gid`; `country`+`city`+`venue` |
| `locations` | venue | `country`+`city`+`venue` |
| `durations` | tagged MP3 track | `gid`; `song` |
| `discography` | release (album/EP) | `releaseid` |
| `songs` | song | `song` (title text); `releaseid` |
| `bands` | show + co-billed band | `gid` |

## Join keys

### `gid` - the show identifier

`gid` is a slug built from city, country, and date (e.g.
`washington-dc-usa-90387`), and is the primary key of `shows`. It is
also the primary key of `durations` and `bands` - join either of those
to `shows` (or to each other) on `gid`.

```{r}
library(fugazi.db)
library(dplyr)

# every show, with its tagged tracks
shows %>%
  left_join(durations, by = "gid")
```

### `country` + `city` + `venue` - venue coordinates

`locations` is keyed by the combination of `country`, `city`, and
`venue` (with `city` disambiguated as `"City (ST)"`/`"City (Country)"` for
towns that share a name with another tour stop - Portland, Columbia,
Croydon, Oxford, Newcastle) - join it onto `shows` on those three
columns to attach coordinates to a show. Coordinates are latitude/longitude
in decimal degrees, using the WGS 84 datum. A handful of venues are
resolved via additional maintainer corrections not present in this table,
so this join won't cover every show.

```{r}
# every show, with its venue's coordinates
shows %>%
  left_join(locations, by = c("country", "city", "venue"))

# venues with no confirmed coordinates yet
locations %>% filter(is.na(longitude) | is.na(latitude))
```

### `song` - song metadata

`songs` is keyed by `song` (the song title, as text), and carries
studio discography metadata (release, vocals, instrumental, duration).
`durations` is also keyed by `song` text for each tagged track - join the two
to attach song metadata to a specific live rendition.

```{r}
# every tagged track, with its song metadata
durations %>%
  left_join(songs, by = "song")
```

### `releaseid` - discography structure

`discography` is keyed by `releaseid`, referenced by `songs$releaseid`.

```{r}
# every song, with its release's name and date
songs %>%
  left_join(discography, by = "releaseid") %>%
  select(song, release, releasedate)
```

This package's data is refreshed periodically by the maintainer as new
shows are documented or corrections are made to the underlying sources.
